---
title: "Reproducible Research (Project 1)"
author: "Oyhx"
date: "29/05/2018"
output: html_document
---

Load Packages
```{r}
library(ggplot2)
library(dplyr)
```


## Loading the Data

```{r import-data}
activity <- read.csv("activity.csv", stringsAsFactors = FALSE)

```


## What is mean total number of steps taken per day?

```{r stepsummarydaily}
dailyactivity <- activity %>%
  group_by(date) %>%
  summarise(tot_steps= sum(steps), mean_steps = mean(steps), med_steps = median(steps))
```

The total, mean and median number of steps is indicated below.

```{r}
print(dailyactivity)
```


## Histogram of the total number of steps taken each day
```{r histstepsummary}
ggplot(data= dailyactivity, aes(x=date, y=tot_steps), na.rm = TRUE) + geom_histogram(stat = "identity") + theme(
  axis.text.x = element_blank())
```

What is the average daily activity pattern?

Make a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)

Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?

```{r stepsummaryinterval}
intervalactivity <- activity %>%
  group_by(interval) %>%
  summarise(tot_intsteps= sum(steps, na.rm = TRUE), mean_intsteps = mean(steps, na.rm = TRUE), med_intsteps = median(steps, na.rm = TRUE))
```

```{r histogramtotaldaily}
ggplot(data= intervalactivity, aes(x=interval, y=mean_intsteps)) +geom_line()
```

The five minute interval with the highest mean number of steps is 0835.

## Calculate and report the total number of missing values in the dataset (i.e. the total number of rows with NAs)

```{r checkNA}
activity %>%
  select(steps, date, interval) %>%  # replace to your needs
  summarise_all(funs(sum(is.na(.))))
```

## Replace NA values using mean value for interval 

```{r}
activity1 <- activity
library(plyr)
```

```{r}
impute.mean <- function (x) replace (x, is.na(x), mean (x, na.rm = TRUE))
activity1 <- ddply(activity1, ~interval, transform, steps = impute.mean(steps))
```


```{r}
detach(package:plyr)
```

## Daily Plot of updated total number of steps

```{r stepsummarydailynew}
dailyactivity1 <- activity1 %>%
  group_by(date) %>%
  summarise(tot_steps1= sum(steps), mean_steps1 = mean(steps), med_steps1 = median(steps))
```


```{r histstepsummarynew}
ggplot(data= dailyactivity1, aes(x=date, y=tot_steps1), na.rm = TRUE) + geom_histogram(stat = "identity") + theme(
  axis.text.x = element_blank())
```


## Comparing mean and median between original and updated dataset (activity vs activity1)

```{r}
mean(activity$steps, na.rm = TRUE) - mean(activity1$steps)
```

```{r}
median(activity$steps, na.rm = TRUE) - median(activity1$steps)
```

There is no difference in mean and median number of steps between both datasets.

## Are there differences in activity patterns between weekdays and weekends?

Convert the date from character to date data type before using the weekday function.
```{r convertdatetype}
activity1$date <- as.Date(activity1$date)
```

```{r weekdayfunction}
activity1$daytype <- weekdays(activity1$date)
```

```{r groupdayofweek}
activity1$grouped <- ifelse((activity1$daytype == "Saturday"| activity1$daytype == "Sunday"), "Weekend", "Weekday")
```

```{r newsummarytable}
meanact1 <- activity1 %>%
  group_by(grouped, interval) %>%
  summarise(avgsteps = mean(steps))
```

```{r weekdayvsweekendplot}
ggplot(data =meanact1, aes(x=interval, y=avgsteps)) +geom_line() +facet_grid(grouped ~.)
```

From the plot above, there is higher activity between 9am to 10 am during weekday compared to weekend. 